---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dpi = 92,
  fig.retina = 2
)
```

# cvms
**Cross-Validation for Model Selection**  

R package: Cross-validating gaussian and binomial regression models.  

By Ludvig R. Olsen and Benjamin Zachariae.  
Started in Oct. 2016 

Contact at:
r-pkgs@ludvigolsen.dk

Main functions:  

* cross_validate()  
* validate()  
* baseline()  
* combine_predictors()  
* cv_plot()  
* select_metrics()  
* reconstruct_formulas()  

## Important News

* AUC calculation has changed. Now explicitly sets the direction in pROC::roc. (27th of May 2019)

* Unit tests have been updated for the new random sampling generator in R 3.6.0. They will NOT run on previous versions of R. 

* Argument "positive" now defaults to 2. If a dependent variable has the values 0 and 1, 1 is now the default positive class, as that's the second largest values. If the dependent variable is a character, it's in alphabetical order.

* Results now contain a count of singular fit messages. See ?lme4::isSingular for more information.

## Installation  

CRAN:

> install.packages("cvms")

Development version:  

> install.packages("devtools")  
>
> devtools::install_github("LudvigOlsen/groupdata2")  
>
> devtools::install_github("LudvigOlsen/cvms")  

# Examples

## Attach packages
```{r warning=FALSE, message=FALSE}
library(cvms)
library(groupdata2) # fold()
library(knitr) # kable()
library(dplyr) # %>% arrange()
library(ggplot2)
```

## Load data

The dataset participant.scores comes with cvms.

```{r}
data <- participant.scores
```

## Fold data

Create a grouping factor for subsetting of folds using groupdata2::fold(). Order the dataset by the folds.

```{r}
# Set seed for reproducibility
set.seed(7)

# Fold data 
data <- fold(data, k = 4,
             cat_col = 'diagnosis',
             id_col = 'participant') %>% 
  arrange(.folds)

# Show first 15 rows of data
data %>% head(15) %>% kable()
```

## Cross-validate a single model

### Gaussian

```{r warning=FALSE, message=FALSE}
CV1 <- cross_validate(data, "score~diagnosis", 
                     fold_cols = '.folds', 
                     family='gaussian', 
                     REML = FALSE)

# Show results
CV1

# Let's take a closer look at the different parts of the output 

# Results metrics
CV1 %>% select_metrics() %>% kable()

# Nested predictions 
# Note that [[1]] picks predictions for the first row
CV1$Predictions[[1]] %>% head() %>% kable()

# Nested results from the different folds
CV1$Results[[1]] %>% kable()

# Nested model coefficients
# Note that you have the full p-values, 
# but kable() only shows a certain number of digits
CV1$Coefficients[[1]] %>% kable()

# Additional information about the model
# and the training process
CV1 %>% select(11:17) %>% kable()

```

### Binomial

```{r}
CV2 <- cross_validate(data, "diagnosis~score", 
                     fold_cols = '.folds', 
                     family='binomial')

# Show results
CV2

# Let's take a closer look at the different parts of the output 
# We won't repeat the parts too similar to those in Gaussian

# Results metrics
CV2 %>% select(1:9) %>% kable()
CV2 %>% select(10:14) %>% kable()

# ROC curve info
CV2$ROC[[1]] %>% head() %>% kable()

# Confusion matrix
CV2$`Confusion Matrix`[[1]] %>% kable()
```


## Cross-validate multiple models

### Create model formulas

```{r}
models <- c("score~diagnosis","score~age")
mixed_models <- c("score~diagnosis+(1|session)","score~age+(1|session)")
```

### Cross-validate fixed effects models

```{r}
CV3 <- cross_validate(data, models, 
                     fold_cols = '.folds', 
                     family='gaussian', 
                     REML = FALSE)

# Show results
CV3
```

### Cross-validate mixed effects models

```{r}
CV4 <- cross_validate(data, mixed_models, 
                     fold_cols = '.folds', 
                     family='gaussian', 
                     REML = FALSE)

# Show results
CV4
```

## Repeated cross-validation

Let's first add some extra fold columns. We will use the num_fold_cols argument to add 3 unique fold columns. We tell fold() to keep the existing fold column and simply add three extra columns. We could also choose to remove the existing fold column, if for instance we were changing the number of folds (k). Note, that the original fold column will be renamed to ".folds_1".  

```{r}
# Set seed for reproducibility
set.seed(2)

# Fold data 
data <- fold(data, k = 4,
             cat_col = 'diagnosis',
             id_col = 'participant',
             num_fold_cols = 3,
             handle_existing_fold_cols = "keep")

# Show first 15 rows of data
data %>% head(10) %>% kable()

```


```{r}
CV5 <- cross_validate(data, "diagnosis ~ score", 
                     fold_cols = paste0(".folds_", 1:4), 
                     family='binomial', 
                     REML = FALSE)

# Show results
CV5

# The binomial output now has a nested 'Results' tibble
# Let's see a subset of the columns
CV5$Results[[1]] %>% select(1:8) %>%  kable()
```

## Baseline evaluations

Create baseline evaluations of a test set.

### Gaussian

Approach: The baseline model (y ~ 1), where 1 is simply the intercept (i.e. mean of y), is fitted on n random subsets of the training set and evaluated on the test set. We also perform an evaluation of the model fitted on the entire training set.

Start by partitioning the dataset.

```{r}
# Set seed for reproducibility
set.seed(1)

# Partition the dataset 
partitions <- groupdata2::partition(participant.scores, 
                                    p = 0.7,
                                    cat_col = 'diagnosis',
                                    id_col = 'participant',
                                    list_out = TRUE)
train_set <- partitions[[1]]
test_set <- partitions[[2]]
```

Create the baseline evaluations.

```{r}
baseline(test_data = test_set, train_data = train_set,
         n = 100, dependent_col = "score", family = "gaussian")
```

### Binomial

Approach: n random sets of predictions are evaluated against the dependent variable in the test set. We also evaluate a set of all 0s and a set of all 1s.

Create the baseline evaluations.

```{r}
baseline(test_data = test_set, n = 100, 
         dependent_col = "diagnosis", family = "binomial")
```


## Plot results

There are currently a small set of plots for quick visualization of the results. It is supposed to be easy to extract the needed information to create your own plots. If you lack access to any information or have other requests or ideas, feel free to open an issue.

### Gaussian

```{r}
cv_plot(CV1, type = "RMSE") +
  theme_bw()
cv_plot(CV1, type = "r2") +
  theme_bw()
cv_plot(CV1, type = "IC") +
  theme_bw()
cv_plot(CV1, type = "coefficients") +
  theme_bw()
```

### Binomial

```{r}
cv_plot(CV2, type = "ROC") +
  theme_bw()
```

## Generate model formulas

Instead of manually typing all possible model formulas for a set of fixed effects (including the possible interactions), combine_predictors() can do it for you (with some constraints). 

When including interactions, >200k formulas have been precomputed for up to 8 fixed effects, with a maximum interaction size of 3, and a maximum of 5 fixed effects per formula. It's possible to further limit the generated formulas.

We can also append a random effects structure to the generated formulas.

```{r}
combine_predictors(dependent = "y",
                   fixed_effects = c("a","b","c"),
                   random_effects = "(1|d)")
```

If two or more fixed effects should not be in the same formula, like an effect and its log-transformed version, we can provide them as sublists.

```{r}
combine_predictors(dependent = "y",
                   fixed_effects = list("a", list("b","log_b")),
                   random_effects = "(1|d)")
```

