---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# cvms
**Cross-Validation for Model Selection**  

R package: Cross-validating gaussian and binomial regression models.  

By Ludvig R. Olsen and Benjamin Zachariae,  
Cognitive Science, Aarhus University.  
Started in Oct. 2016 

Contact at:
r-pkgs@ludvigolsen.dk

Main functions:  

* cross_validate()
* validate()
* cv_plot()

## Installation  

Development version:  

> install.packages("devtools")  
>
> devtools::install_github("LudvigOlsen/groupdata2")  
>
> devtools::install_github("LudvigOlsen/cvms")  


# Examples

## Attach packages
```{r warning=FALSE, message=FALSE}
library(cvms)
library(groupdata2) # fold()
library(knitr) # kable()
library(dplyr) # %>% arrange()
library(ggplot2)
```

## Load data

The dataset participant.scores comes with cvms.

```{r}
data <- participant.scores
```

## Fold data

Create a grouping factor for subsetting in folds using groupdata2::fold(). Order the dataset by the folds.

```{r}
# Set seed for reproducibility
set.seed(7)

# Fold data 
data <- fold(data, k = 4,
             cat_col = 'diagnosis',
             id_col = 'participant') %>% 
  arrange(.folds)

# Show first 15 rows of data
data %>% head(15) %>% kable()
```

## Cross-validate a single model

### Gaussian

```{r}
CV1 <- cross_validate(data, "score~diagnosis", 
                     fold_cols = '.folds', 
                     family='gaussian', 
                     REML = FALSE)

# Show results
CV1

# Let's take a closer look at the different parts of the output 

# Results metrics
CV1 %>% select(1:7) %>% kable()

# Nested predictions 
# Note that [[1]] picks predictions for the first row
CV1$Predictions[[1]] %>% head() %>% kable()

# Nested results from the different folds/models
CV1$Results[[1]] %>% kable()

# Nested model coefficients
# Note that you have the full p-values, 
# but kable() only show a certain number of digits
CV1$Coefficients[[1]] %>% kable()

# Additional information about the model
# and the training process
CV1 %>% select(11:17) %>% kable()

```

### Binomial

```{r}
CV2 <- cross_validate(data, "diagnosis~score", 
                     fold_cols = '.folds', 
                     family='binomial')

# Show results
CV2

# Let's take a closer look at the different parts of the output 
# We won't repeat the parts too similar to those in Gaussian

# Results metrics
CV2 %>% select(1:9) %>% kable()
CV2 %>% select(10:14) %>% kable()

# ROC curve info
CV2$ROC[[1]] %>% head() %>% kable()
```


## Cross-validate multiple models

### Create model formulas

```{r}
models <- c("score~diagnosis","score~age")
mixed_models <- c("score~diagnosis+(1|session)","score~age+(1|session)")
```

### Cross-validate fixed effects models

```{r}
CV3 <- cross_validate(data, models, 
                     fold_cols = '.folds', 
                     family='gaussian', 
                     REML = FALSE)

# Show results
CV3
```

### Cross-validate mixed effects models

```{r}
CV4 <- cross_validate(data, mixed_models, 
                     fold_cols = '.folds', 
                     family='gaussian', 
                     REML = FALSE)

# Show results
CV4
```

## Repeated cross-validation
Note: currently only work with the github version of groupdata2!

Let's first create new folds. We will use the num_fold_cols argument to add 3 unique fold columns.

```{r}
# devtools::install_github("ludvigolsen/groupdata2", ref="repeatedCV")

# Set seed for reproducibility
set.seed(7)

# Fold data 
data <- fold(data, k = 4,
             cat_col = 'diagnosis',
             id_col = 'participant',
             num_fold_cols = 3)

# Show first 15 rows of data
data %>% head(10) %>% kable()

```


```{r}
CV5 <- cross_validate(data, "diagnosis~score+(1|session)", 
                     fold_cols = c('.folds_1','.folds_2','.folds_3'), 
                     family='binomial', 
                     REML = FALSE)

# Show results
CV5

# The binomial output now has a nested results tibble
# Let's see a subset of the columns
CV5$Results[[1]] %>% select(1:8) %>%  kable()
```


## Plot results

There are currently a small set of plots for quick visualization of the results. It is supposed to be easy to extract the needed information to create your own plots. If you lack access to any information or have other requests or ideas, feel free to open an issue.

### Gaussian

```{r}
cv_plot(CV1, type = "RMSE") +
  theme_bw()
cv_plot(CV1, type = "r2") +
  theme_bw()
cv_plot(CV1, type = "IC") +
  theme_bw()
cv_plot(CV1, type = "coefficients") +
  theme_bw()
```

### Binomial

```{r}
cv_plot(CV2, type = "ROC") +
  theme_bw()
```


