---
title: "Introduction to cvms"
author: 
  - "Ludvig Renbo Olsen"
date: "`r Sys.Date()`"
abstract: |
  This vignette is an introduction to the package cvms.  
  Cross-Validation for Model Selection.  
  &nbsp;  
  &nbsp;  
  Contact author at r-pkgs@ludvigolsen.dk
  &nbsp;  
  &nbsp;  
  
output: 
  rmarkdown::html_vignette:
    css: 
    - !expr system.file("rmarkdown/templates/html_vignette/resources/vignette.css", package = "rmarkdown")
    - styles.css
    fig_width: 6
    fig_height: 4
    toc: yes
    number_sections: no
  rmarkdown::pdf_document:
    highlight: tango
    number_sections: yes
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Introduction_to_cvms}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/vignette_intro-",
  dpi = 92,
  fig.retina = 2
)
```

# Main functions

The main functions of cvms are:  

* `cross_validate()`  
* `validate()`  
* `evaluate()`
* `baseline()`  
* `combine_predictors()`  
* `cv_plot()`  
* `select_metrics()`  
* `reconstruct_formulas()` 

# Examples

## Attach packages
```{r warning=FALSE, message=FALSE}
library(cvms)
library(groupdata2) # fold() partition()
library(knitr) # kable()
library(dplyr) # %>% arrange()
library(ggplot2)
```

## Load data

The dataset `participant.scores` comes with cvms.

```{r}
data <- participant.scores
```

## Fold data

Create a grouping factor for subsetting of folds using `groupdata2::fold()`. Order the dataset by the folds.

```{r}
# Set seed for reproducibility
set.seed(7)

# Fold data 
data <- fold(data, k = 4,
             cat_col = 'diagnosis',
             id_col = 'participant') %>% 
  arrange(.folds)

# Show first 15 rows of data
data %>% head(15) %>% kable()
```

## Cross-validate a single model

### Gaussian

```{r warning=FALSE, message=FALSE}
CV1 <- cross_validate(data, "score~diagnosis",
                      fold_cols = '.folds',
                      family = 'gaussian',
                      REML = FALSE)

# Show results
CV1

# Let's take a closer look at the different parts of the output 

# Results metrics
CV1 %>% select_metrics() %>% kable()

# Nested predictions 
# Note that [[1]] picks predictions for the first row
CV1$Predictions[[1]] %>% head() %>% kable()

# Nested results from the different folds
CV1$Results[[1]] %>% kable()

# Nested model coefficients
# Note that you have the full p-values, 
# but kable() only shows a certain number of digits
CV1$Coefficients[[1]] %>% kable()

# Additional information about the model
# and the training process
CV1 %>% select(11:17) %>% kable()

```

### Binomial

```{r}
CV2 <- cross_validate(data, "diagnosis~score",
                      fold_cols = '.folds',
                      family = 'binomial')

# Show results
CV2

# Let's take a closer look at the different parts of the output 
# We won't repeat the parts too similar to those in Gaussian

# Results metrics
CV2 %>% select(1:9) %>% kable()
CV2 %>% select(10:14) %>% kable()

# ROC curve info
CV2$ROC[[1]] %>% head() %>% kable()

# Confusion matrix
CV2$`Confusion Matrix`[[1]] %>% kable()
```


## Cross-validate multiple models

### Create model formulas

```{r}
models <- c("score~diagnosis", "score~age")
mixed_models <- c("score~diagnosis+(1|session)", "score~age+(1|session)")
```

### Cross-validate fixed effects models

```{r}
CV3 <- cross_validate(data, models,
                      fold_cols = '.folds',
                      family = 'gaussian',
                      REML = FALSE)

# Show results
CV3
```

### Cross-validate mixed effects models

```{r}
CV4 <- cross_validate(data, mixed_models,
                      fold_cols = '.folds',
                      family = 'gaussian',
                      REML = FALSE)

# Show results
CV4
```

## Repeated cross-validation

Let's first add some extra fold columns. We will use the num_fold_cols argument to add 3 unique fold columns. We tell `fold()` to keep the existing fold column and simply add three extra columns. We could also choose to remove the existing fold column, if for instance we were changing the number of folds (k). Note, that the original fold column will be renamed to ".folds_1".  

```{r}
# Set seed for reproducibility
set.seed(2)

# Fold data 
data <- fold(data, k = 4,
             cat_col = 'diagnosis',
             id_col = 'participant',
             num_fold_cols = 3,
             handle_existing_fold_cols = "keep")

# Show first 15 rows of data
data %>% head(10) %>% kable()

```


```{r}
CV5 <- cross_validate(data, "diagnosis ~ score",
                      fold_cols = paste0(".folds_", 1:4),
                      family = 'binomial',
                      REML = FALSE)

# Show results
CV5

# The binomial output now has a nested 'Results' tibble
# Let's see a subset of the columns
CV5$Results[[1]] %>% select(1:8) %>%  kable()
```

## Evaluating predictions

Evaluate predictions from a model trained outside cvms. Works with linear regression (`gaussian`), logistic regression (`binomial`), and multiclass classification (`multinomial`). The following is an example of multinomial evaluation.

### Multinomial

Create a dataset with 3 predictors and a target column. Partition it with `groupdata2::partition()` to create a training set and a validation set. `multiclass_probability_tibble()` is a simple helper function for generating random tibbles.

```{r}
# Set seed
set.seed(1)

# Create class names
class_names <- paste0("class_", 1:4)

# Create random dataset with 100 observations 
# Partition into training set (75%) and test set (25%)
multiclass_partitions <- multiclass_probability_tibble(
  num_classes = 3, # Here, number of predictors
  num_observations = 100,
  apply_softmax = FALSE,
  FUN = rnorm,
  class_name = "predictor_") %>%
  dplyr::mutate(class = sample(
    class_names,
    size = 100,
    replace = TRUE)) %>%
  partition(p = 0.75,
            cat_col = "class")

# Extract partitions
multiclass_train_set <- multiclass_partitions[[1]]
multiclass_test_set <- multiclass_partitions[[2]]

multiclass_test_set
```

Train multinomial model using the `nnet` package and get the predicted probabilities. 

```{r}
# Train multinomial model
multiclass_model <- nnet::multinom(
   "class ~ predictor_1 + predictor_2 + predictor_3",
   data = multiclass_train_set)

# Predict the targets in the test set
predictions <- predict(multiclass_model, 
                       multiclass_test_set,
                       type = "probs") %>%
  dplyr::as_tibble()

# Add the targets
predictions[["target"]] <- multiclass_test_set[["class"]]

head(predictions, 10)
```

Perform the evaluation. This will create one-vs-all binomial evaluations and summarize the results.

```{r}
# Evaluate predictions
evaluate(data = predictions,
         target_col = "target",
         prediction_cols = class_names,
         type = "multinomial")
```


## Baseline evaluations

Create baseline evaluations of a test set.

### Gaussian

Approach: The baseline model (y ~ 1), where 1 is simply the intercept (i.e. mean of y), is fitted on n random subsets of the training set and evaluated on the test set. We also perform an evaluation of the model fitted on the entire training set.

Start by partitioning the dataset.

```{r}
# Set seed for reproducibility
set.seed(1)

# Partition the dataset 
partitions <- groupdata2::partition(participant.scores,
                                    p = 0.7,
                                    cat_col = 'diagnosis',
                                    id_col = 'participant',
                                    list_out = TRUE)
train_set <- partitions[[1]]
test_set <- partitions[[2]]
```

Create the baseline evaluations:

```{r}
baseline(test_data = test_set, train_data = train_set,
         n = 100, dependent_col = "score", family = "gaussian")
```

### Binomial

Approach: n random sets of predictions are evaluated against the dependent variable in the test set. We also evaluate a set of all 0s and a set of all 1s.

Create the baseline evaluations:

```{r}
baseline(test_data = test_set, n = 100, 
         dependent_col = "diagnosis", family = "binomial")
```

### Multinomial

Approach: Creates one-vs-all (binomial) baseline evaluations for n sets of random predictions against the dependent variable, along with sets of "all class x,y,z,..." predictions.

Create the baseline evaluations:

```{r}
multiclass_baseline <- baseline(
  test_data = multiclass_test_set, n = 100,
  dependent_col = "class", family = "multinomial")

# Summarized metrics
multiclass_baseline$summarized_metrics

# Summarized class level results for class 1
multiclass_baseline$summarized_class_level_results %>% 
  dplyr::filter(Class == "class_1") %>%
  tidyr::unnest(Results)

# Random evaluations
# Note, that the class level results for each repetition
# is available as well
multiclass_baseline$random_evaluations
```

## Plot results

There are currently a small set of plots for quick visualization of the results. It is supposed to be easy to extract the needed information to create your own plots. If you lack access to any information or have other requests or ideas, feel free to open an issue.

### Gaussian

```{r}
cv_plot(CV1, type = "RMSE") +
  theme_bw()
cv_plot(CV1, type = "r2") +
  theme_bw()
cv_plot(CV1, type = "IC") +
  theme_bw()
cv_plot(CV1, type = "coefficients") +
  theme_bw()
```

### Binomial

```{r}
cv_plot(CV2, type = "ROC") +
  theme_bw()
```

## Generate model formulas

Instead of manually typing all possible model formulas for a set of fixed effects (including the possible interactions), `combine_predictors()` can do it for you (with some constraints). 

When including interactions, >200k formulas have been precomputed for up to 8 fixed effects, with a maximum interaction size of 3, and a maximum of 5 fixed effects per formula. It's possible to further limit the generated formulas.

We can also append a random effects structure to the generated formulas.

```{r}
combine_predictors(dependent = "y",
                   fixed_effects = c("a","b","c"),
                   random_effects = "(1|d)")
```

If two or more fixed effects should not be in the same formula, like an effect and its log-transformed version, we can provide them as sublists.

```{r}
combine_predictors(dependent = "y",
                   fixed_effects = list("a", list("b","log_b")),
                   random_effects = "(1|d)")
```

